---
title: "Assignment 2"
author: "Arnav Murulidhar"
date: "9/14/2020"
output: html_document
---

## Setup

The libraries needed for the assignment:

```{r load libraries, message=FALSE}
library(tidyverse)
library(ggplot2)
```

Then I'll load the dataset from Assignment 1 and attach it to the search path

```{r load and attach data, message=FALSE}
tract_data <- read_csv("commutersLosAngelesCA2018.csv")
attach(tract_data)
```

### List of variables (from Assignment 1):

1. Property value (continuous) : VALP

2. Units in structure (categorical) : BLD

3. Total person's income (continuous) : PINCP

4. Means of transportation to work (categorical) : JWTR

5. Travel time to work (continuous) : JWMNP

## Distribution of Continuous Variables

### Descriptive statistics

First I calculate summary statistics.

```{r}
prop_summary <-  summary(VALP)
prop_sd <- sd(VALP, na.rm = TRUE)
prop_mean_int <- t.test(VALP)
income_summary <- summary(PINCP)
income_sd <- sd(PINCP, na.rm = TRUE)
income_mean_int <- t.test(PINCP)
commute_summary <- summary(JWMNP)
commute_sd <- sd(JWMNP, na.rm = TRUE)
commute_mean_int <- t.test(JWMNP)
```

And display them in a formatted table:

```{r}
summary_table <- tibble(
  Statistic = c("Sample mean", 
                "Median",
                "Standard deviation",
                "Interquartile range",
                "Population mean (95% confidence)"),
  `Property value ( $ )` = c(prettyNum(prop_summary["Mean"], digits=3),
                           prop_summary["Median"],
                           prettyNum(prop_sd, digits = 3),
                           paste(prop_summary["1st Qu."],"to", 
                                 prop_summary["3rd Qu."]),
                           paste(prettyNum(prop_mean_int$conf.int[1],
                                           digits=3), "to",
                                 prettyNum(prop_mean_int$conf.int[2],
                                               digits=3))),
  `Income ( $ )` = c(paste(prettyNum(income_summary["Mean"], digits=3)),
                           paste(prettyNum(income_summary["Median"],digits=3)),
                           paste(prettyNum(income_sd, digits=3)),
                           paste(prettyNum(income_summary["1st Qu."], digits=3),
                                 " to", 
                                 prettyNum(income_summary["3rd Qu."], digits=3)),
                           paste(prettyNum(income_mean_int$conf.int[1],
                                           digits=3), " to",
                                 prettyNum(income_mean_int$conf.int[2],
                                           digits=3))),
  `Commute time (min)` = c(paste(prettyNum(commute_summary["Mean"], digits=3)),
                          paste(prettyNum(commute_summary["Median"], digits=3)),
                          paste(prettyNum(commute_sd, digits=3)),
                          paste(prettyNum(commute_summary["1st Qu."], digits=3), 
                                " to", 
                          prettyNum(commute_summary["3rd Qu."], digits=3)),
                          paste(prettyNum(commute_mean_int$conf.int[1], 
                                           digits=3), " to",
                          prettyNum(commute_mean_int$conf.int[2], 
                                           digits=3))))
knitr::kable(summary_table, 
             caption = "Characteristics of census tracts in Los Angeles County")
```

### Histograms

Now, I'll create histogram for each of these three distributions. 

#### Property Value

```{r}
ggplot(tract_data, aes(x = VALP)) + 
  geom_histogram(color = "gray", binwidth = 100000) +
  theme_bw() +
  scale_x_continuous(name = "Median property value in tract",
                     breaks = breaks <- seq(0, 6000000, by = 1000000)) + 
  scale_y_continuous(name = "Number of properties in sample") 
```

The distribution of property value looks positively skewed, thus the median value is less than the corresponding mean value. The outlier values increases the mean quite a bit.

#### Income 

```{r}
ggplot(tract_data, aes(x = PINCP)) + 
  geom_histogram(color = "gray", binwidth = 10000) +
  theme_bw() +
  scale_x_continuous(name = "Income",
                     breaks = breaks <- seq(0, 1000000, by = 100000)) + 
  scale_y_continuous(name = "Number of people in sample") 
```

The distribution of income is also skewed to the right, with a long tail.

#### Commute time


```{r}
ggplot(tract_data, aes(x = JWMNP)) + 
  geom_histogram(color = "gray", binwidth = 10) +
  theme_bw() +
  scale_x_continuous(name = 
    "Commute time",
                     breaks = breaks <- seq(0, 140, by = 20)) + 
  scale_y_continuous(name = "Number of people in sample") 
```

The distribution of commute time is slightly skewed to the right, with outliers at the 140min mark. However, you can see that the bulk of commute times range between 15-45 minutes.

## Distribution of Categorical Variables

I'll show two tables, one showing the proportions of each category for each of my categorical variables.

### Unit types

First I'll do a t-test for each housing type.

```{r}

one_det <- t.test (BLD == "02")
three_apt <- t.test (BLD == "05")
one_att <- t.test (BLD == "03")
ten_apt <- t.test (BLD == "07")
mob_home <- t.test (BLD == "01")
twen_apt <- t.test (BLD == "08")
fifty_apt <- t.test (BLD == "09")
five_apt <- t.test (BLD == "06")
two_apt <- t.test (BLD == "04")
boat_rv <- t.test (BLD == "10")
```

Then I'll create a formatted table with proportions and confidence intervals

```{r}
race_shares <- tibble(Majority_Race = c("One-family house detached",
                                  "3-4 Apartments",
                                  "One-family house attached",
                                  "10-19 Apartments",
                                  "Mobile home or trailer",
                                  "20-49 Apartments",
                                  "50 or more Apartments",
                                  "5-9 Apartments",
                                  "2 Apartments",
                                  "Boat, RV, van, etc."),
                 Sample_proportion = c(paste(prettyNum(one_det$estimate*100, 
                                                         digits = 3), "%"),
                                         paste(prettyNum(three_apt$estimate*100, 
                                                         digits = 3), "%"),
                                         paste(prettyNum(one_att$estimate*100, 
                                                         digits = 3), "%"),
                                         paste(prettyNum(ten_apt$estimate*100, 
                                                         digits = 3), "%"),
                                         paste(prettyNum(mob_home$estimate*100, 
                                                         digits = 3), "%"),
                                         paste(prettyNum(twen_apt$estimate*100, 
                                                         digits = 3), "%"),
                                         paste(prettyNum(fifty_apt$estimate*100, 
                                                         digits = 3), "%"),
                                         paste(prettyNum(five_apt$estimate*100, 
                                                         digits = 3), "%"),
                                         paste(prettyNum(two_apt$estimate*100, 
                                                         digits = 3), "%"),
                                         paste(prettyNum(boat_rv$estimate*100, 
                                                         digits = 3), "%")),
                 `95-percent confidence interval` = c(
                   paste(prettyNum(one_det$conf.int[1]*100, digits=3),
                         "% to ",
                         prettyNum(one_det$conf.int[2]*100, digits=3),
                         "%", sep=""),
                   paste(prettyNum(three_apt$conf.int[1]*100, digits=3),
                         "% to ",
                         prettyNum(three_apt$conf.int[2]*100, digits=3),
                         "%", sep=""),
                   paste(prettyNum(one_att$conf.int[1]*100, digits=3),
                         "% to ",
                         prettyNum(one_att$conf.int[2]*100, digits=3),
                         "%", sep=""),
                   paste(prettyNum(ten_apt$conf.int[1]*100, digits=3),
                         "% to ",
                         prettyNum(ten_apt$conf.int[2]*100, digits=3),
                         "%", sep=""),
                   paste(prettyNum(mob_home$conf.int[1]*100, digits=3),
                         "% to ",
                         prettyNum(mob_home$conf.int[2]*100, digits=3),
                         "%", sep=""),
                   paste(prettyNum(twen_apt$conf.int[1]*100, digits=3),
                         "% to ",
                         prettyNum(twen_apt$conf.int[2]*100, digits=3),
                         "%", sep=""),
                   paste(prettyNum(fifty_apt$conf.int[1]*100, digits=3),
                         "% to ",
                         prettyNum(fifty_apt$conf.int[2]*100, digits=3),
                         "%", sep=""),
                   paste(prettyNum(five_apt$conf.int[1]*100, digits=3),
                         "% to ",
                         prettyNum(five_apt$conf.int[2]*100, digits=3),
                         "%", sep=""),
                   paste(prettyNum(two_apt$conf.int[1]*100, digits=3),
                         "% to ",
                         prettyNum(two_apt$conf.int[2]*100, digits=3),
                         "%", sep=""),
                   paste(prettyNum(boat_rv$conf.int[1]*100, digits=3),
                         "% to ",
                         prettyNum(boat_rv$conf.int[2]*100, digits=3),
                         "%", sep="")))
                   
                   
knitr::kable(race_shares, caption = "Census tracts in Los Angeles County")
```






